---
title: "MOVES Model Interpolation"
author: "Cordero Ortiz -- SUPR Lab"
date: "8/9/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

# Introduction 

This document will detail how outputs from diesel particulate matter MOVES model were interpolated using Ordinary Krigging.

# Packages Used

```{r}
library(sp)
library(sf)
library(raster)
library(gstat)
library(dplyr)
```

# MOVES Model Outputs

Model data were provided with DAT file extensions -- one for the current inventory of construction vehicles and another for the Tier 4 vehicle scenario. We can take a peek at the first few lines to determine what format the data within the .DAT file is in:
```{r}
readLines("./data/diesel_moves_output/CurrentInventory_DSLPM_8760HR_CONC.DAT", n = 10)
```
 
We now know that data begins on line 7 of the DAT files. Working with the current inventory file:
```{r}
currentInv <- read.table("./data/diesel_moves_output/CurrentInventory_DSLPM_8760HR_CONC.DAT", header = FALSE, skip = 6)
str(currentInv)
```

Variables V1&2 are receptor coordinates in WGS84/UTM Zone 10N projected coordinate system while variables coordinates in WGS84. Variable V3 is the modeled DPM in ug/m^3 or ppb.
Let's stick to the projected coordinate system and grab the DPM from the data frame.

```{r}
coords <- currentInv[, 1:2] #grab UTM Zone 10N Km
coords <- apply(coords, MARGIN = c(1,2), FUN = function(x){x*1000}) # now in meters

data <- currentInv[, 3] %>% as.data.frame() # grab data 
currentInv <- cbind(data, coords)
#dat <- dat[,3:ncol(dat)]
names(currentInv) <- c("dpm_ppm", "lat", "lon")
str(currentInv)
```

Repeating the above process for the Tier 4 Inventory Scenario.
```{r}
tier4Inv <- read.table("./data/diesel_moves_output/Tier4_Scenario_DSLPM_8760HR_CONC.DAT", header = FALSE, skip = 6)
coords <- tier4Inv[, 1:2] #grab UTM Zone 10N Km
coords <- apply(coords, MARGIN = c(1,2), FUN = function(x){x*1000}) # now in meters

data <- tier4Inv[, 3] %>% as.data.frame() # grab data 
tier4Inv <- cbind(data, coords)
#dat <- dat[,3:ncol(dat)]
names(tier4Inv) <- c("dpm_ppm", "lat", "lon")
str(tier4Inv)
```

# Kriging

## Fitting the Variogram Model

The first step in krigging is the fit a variogram model to the data. The `gstat` package accepts `sp` objects, so we will first create an `sp` obejct using the current inventory model: 
```{r}
coordinates(currentInv) <- ~lat + lon # build `sp` object
# assign CRS to `sp` object
currentInv@proj4string <- CRS(as.character("+proj=utm +zone=10 +ellps=WGS84 +datum=WGS84 +units=m +no_defs")) #

# 1st order polynomial equation
# f.1 <- as.formula(dpm_ppm ~ X + Y) # Universal
f.1 <- as.formula(dpm_ppm ~ 1) # Ordinary


# # Add X/Y coordinates to the `sp` data frame
# currentInv$X <- coordinates(currentInv)[, 1]
# currentInv$Y <-  coordinates(currentInv)[, 2]


# Create variogram
var.smpl <- variogram(object = f.1, data = currentInv, cloud = FALSE)

currentInv.fit <- fit.variogram(var.smpl,
                        vgm(psill=10, model="Sph", range=30000, nugget=10))

plot(var.smpl, currentInv.fit, xlim=c(0, 40000), ylim = c(0, 40))
```

## Surface Generation

### Create Empty Grid

We need a grid to create a krigged surface, so we will create one using our point data:



```{r}
## What cell size?

require(units)
dist[dist==as_units(0, "m")] <- NA
NND <- apply(dist, 2, min,na.rm=TRUE) %>% round(2)
mean(NND, na.rm=TRUE) # 1Km grid seems appropriate 

grid <- sp::makegrid(currentInv, cellsize = 1000)
names(grid) <- c("X", "Y")
coordinates(grid) <- c("X", "Y")
gridded(grid) <- TRUE  # Create SpatialPixel object
fullgrid(grid) <- TRUE  # Create SpatialGrid object


# Add P's projection information to the empty grid
grid@proj4string <- CRS(as.character("+proj=utm +zone=10 +ellps=WGS84 +datum=WGS84 +units=m +no_defs"))

plot1 <- tier4Inv %>% as.data.frame %>% 
  ggplot(aes(lon, lat)) + geom_point(size=1) + 
  ggtitle("Points with Measurements") + coord_equal()

plot2 <- grid %>% as.data.frame() %>% 
  ggplot(aes(X, Y)) + geom_point(size=1) + coord_equal() + 
  ggtitle("Points to Estimate")

library(gridExtra)
grid.arrange(plot1, plot2, ncol = 2)

```


### Krigging

#### Current Inventory
```{r}
currentInv <- currentInv[-zerodist(currentInv)[,1],]  # Remove repeated Points
sp::zerodist(currentInv)[, 1]

# f.1 <- as.formula(dpm_ppm ~ X + Y)
f.1 <- as.formula(dpm_ppm ~ 1)

# Perform the krige interpolation (note the use of the variogram model
# created in the earlier step)
currentInv.krg <- krige( f.1, currentInv, grid, currentInv.fit)

currentInv.krg.r <- raster::raster(currentInv.krg, layer="var1.pred")

metro <- st_read("./data/location_shapefiles/metro.shp") %>% st_transform(32610)

currentInv.krg.r.m <- raster::mask(currentInv.krg.r, metro) # clip surface to pdx metro area

qtm(currentInv.krg.r.m)

```

#### Tier 4 Scenario
```{r}

coordinates(tier4Inv) <- ~lat + lon # build `sp` object
# assign CRS to `sp` object
tier4Inv@proj4string <- CRS(as.character("+proj=utm +zone=10 +ellps=WGS84 +datum=WGS84 +units=m +no_defs")) #

# 1st order polynomial equation
# f.1 <- as.formula(dpm_ppm ~ X + Y) # Universal
f.1 <- as.formula(dpm_ppm ~ 1) # Ordinary


# # Add X/Y coordinates to the `sp` data frame
# tier4Inv$X <- coordinates(tier4Inv)[, 1]
# tier4Inv$Y <-  coordinates(tier4Inv)[, 2]


# Create variogram
var.smpl <- variogram(object = f.1, data = tier4Inv, cloud = FALSE)

tier4Inv.fit <- fit.variogram(var.smpl,
                        vgm(psill=10, model="Sph", range=30000, nugget=10))

plot(var.smpl, tier4Inv.fit, xlim=c(0, 40000), ylim = c(0, 40))


tier4Inv <- tier4Inv[-zerodist(tier4Inv)[,1],]  # Remove repeated Points
sp::zerodist(tier4Inv)[, 1]

# Perform the krige interpolation (note the use of the variogram model
# created in the earlier step)
tier4Inv.krg <- krige( f.1, tier4Inv, grid, tier4Inv.fit)

tier4Inv.krg.r <- raster::raster(tier4Inv.krg, layer="var1.pred")

tier4Inv.krg.r.m <- raster::mask(tier4Inv.krg.r, metro) # clip surface to pdx metro area

qtm(tier4Inv.krg.r.m, ) + qtm(currentInv.krg.r.m)

tier4Inv.krg.r.m@legend
```

# Testing Surfaces
```{r}
test <- (currentInv.krg.r.m / tier4Inv.krg.r.m)

names(currentInv.krg.r.m) <- "Current"
names(tier4Inv.krg.r.m) <- "Tier4"
percent.delta <- (currentInv.krg.r.m / tier4Inv.krg.r.m)
names(percent.delta) <- "times.above.bench"

qtm(tier4Inv.krg.r.m, ) + qtm(currentInv.krg.r.m) + qtm(percent.delta)

test <- test * 100
test %>% qtm()
test <- (tier4Inv.)
qtm(test)



```

# Exposure at locations
```{r}
# lower field names for postgres
lowerName <- function(x){names(x) <- tolower(names(x)); return(x)}

bus_stops <- st_read("./data/location_shapefiles/busstops.shp") %>% st_transform(32610) %>% lowerName() %>% .[metro,]
community_centers <- st_read("./data/location_shapefiles/community_centers.shp") %>% st_transform(32610) %>% lowerName() %>% .[metro,]
hospitals <- st_read("./data/location_shapefiles/hospital.shp") %>% st_transform(32610) %>% lowerName() %>% .[metro,]
libraries <- st_read("./data/location_shapefiles/library.shp") %>% st_transform(32610) %>% lowerName() %>% .[metro,]
lrt_stops <- st_read("./data/location_shapefiles/lrt_stop.shp") %>% st_transform(32610) %>% lowerName() %>% .[metro,]
schools <- st_read("./data/location_shapefiles/schools.shp") %>% st_transform(32610) %>% lowerName() %>% .[metro,]
school_site <- st_read("./data/location_shapefiles/school_site.shp") %>% st_transform(32610) %>% lowerName() %>% .[metro,]
transit_center <- st_read("./data/location_shapefiles/tran_cen.shp") %>% st_transform(32610) %>% lowerName() %>% .[metro,]
```

## Write Rasters and Shapefiles into postgres database
```{r}
library(DBI)
library(RPostgres)
library(rpostgis)

conn <- DBI::dbConnect(RPostgres::Postgres(), 
               dbname = 'diesel2021', 
               host='localhost',
               port='5432',
               user="cordero",
               password='Mojo')

st_write(bus_stops, dsn = conn, layer = "bus_stops", append = FALSE)
st_write(community_centers, dsn = conn, layer = "community_centers", append = FALSE)
st_write(hospitals, dsn = conn, layer = "hospitals", append = FALSE)
st_write(libraries, dsn = conn, layer = "libraries", append = FALSE)
st_write(lrt_stops, dsn = conn, layer = "lrt_stops", append = FALSE)
st_write(schools, dsn = conn, layer = "schools", append = FALSE)
st_write(school_site, dsn = conn, layer = "school_site", append = FALSE)
st_write(transit_center, dsn = conn, layer = "transit_center", append = FALSE)
pgWriteRast(con, raster = currentInv.krg.r.m, name = 'current_scenario')
pgWriteRast(con, raster = tier4Inv.krg.r.m, name = 'tier4_scenario', )



tier4 <- rpostgis::pgGetRast(conn, 'tier4_scenario')



names(vectors)

test <- st_read(con, layer = "schools")
test %>% qtm()
qtm(metro) + qtm(schools)
query <- paste(
  'SELECT *', 
  'FROM schools', 
  'WHERE "TYPE" = \'Public\''
)

query <- paste('
  SELECT s.*, st_value(t4.rast, s.geometry) as "RasterValue", s.geometry
  FROM tier4_scenario AS t4, schools AS s
  WHERE st_intersects(t4.rast, s.geometry)
')

table <- 'schools'
## Polygonizes the raster before intersect
query <- paste('
  SELECT p.*, st_value(t4.rast, p.geometry) as "RasterValue", p.geometry
  FROM tier4_scenario AS t4,', table, ' AS p
  WHERE st_intersects(st_polygon(t4.rast), p.geometry)
')

query <- paste('
  SELECT p.*, st_value(t4.rast, p.geometry) as "RasterValue", p.geometry
  FROM tier4_scenario AS t4, schools AS p
  WHERE (st_intersects(t4.rast, p.geometry) AND type = \'Private\')
  ')

query <- paste('
  SELECT st_convexhull(r.rast)
  FROM tier4_scenario as r;
')

test <- st_read(conn, query=query)

## read a raster 
query <- 'SELECT r.rast FROM tier4_scenario as r;'


?raster::raster()

st_read(conn, query=query)


qtm(tier4Inv.krg.r.m) + qtm(test[!is.na(test$RasterValue),], symbols.col = "RasterValue")


tmap_mode('view')
tm_shape(test[!is.na(test$RasterValue),]) + 
  tm_bubbles(col = "RasterValue", size = 0.1)


library(leaflet)
pal <- colorNumeric(
  palette = "YlOrRd",
  domain = test$RasterValue
)

tier4_pal <- colorNumeric("YlOrRd", values(tier4),
                    na.color = "transparent")
 
test <- test %>% 
  filter(!is.na(RasterValue)) %>% 
  mutate(
    size = ifelse(RasterValue < 0.5, 0.5, RasterValue)
  )

leaflet(test[!is.na(test$RasterValue),] %>% st_transform(4326), elementId = 'map') %>% 
  addProviderTiles(provider = "CartoDB.DarkMatter") %>% 
  addPolygons(data = st_transform(metro, 4326), color = "#b3b3b3", 
              fillOpacity = 0.1, weight = 2, 
              popup = "Portland Metro Area", group = 'pdx') %>% 
  addCircleMarkers(radius = ~5*size, fillColor = ~pal(RasterValue), group = 'points',
                   fillOpacity = 0.7, stroke = FALSE, 
                   clusterOptions = markerClusterOptions(), 
                   clusterId = "pointCluster", 
                   popup = ~paste("<b>School Name:</b>", name, 
                                  "<br><b>DPM Exposure:</b>", round(RasterValue, 2), "ppb",
                                  "<br><b>School Level:</b>", level_name,
                                  "<br><b>School Type:</b>", type,
                                  "<br><b>School Address:</b>", address, ",", state, zipcode)) %>% 
  addRasterImage(tier4, colors = tier4_pal, opacity = 0.5, group = 'tier 4 krigged') %>%
  hideGroup("tier 4 krigged") %>% 
  addLayersControl(overlayGroups = c("pdx", "points", "tier 4 krigged")) %>% 
  addLegend(pal = tier4_pal, values = values(tier4), title = "Tier 4 Scenario", position = "bottomright") %>% 
  addLegend(pal = pal, values = ~RasterValue, position = "bottomright", title = "Diesel PM (ppb)") %>% 
  addEasyButton(easyButton(
    states = list(
      easyButtonState(
        stateName="clustered-markers",
        icon="ion-toggle-filled",
        title="Remove Clusters",
        onClick = JS("
                     function(btn, map) {
                     var clusterManager =
                     map.layerManager.getLayer('cluster', 'pointCluster');
                     clusterManager.disableClustering();
                     btn.state('markers');
                     }")
        ),
  
      easyButtonState(
        stateName="markers",
        icon="ion-toggle",
        title="Enable Clusters",
        onClick = JS("
                     function(btn, map) {
                     var clusterManager =
                     map.layerManager.getLayer('cluster', 'pointCluster');
                     clusterManager.enableClustering();
                     btn.state('clustered-markers');
                     }")
         )
  )
  )) %>% 
  addEasyButton(easyButton(
    icon="fa-crosshairs", title="Locate Me",
    onClick=JS("function(btn, map){ map.locate({setView: true, maxZoom: 15}); }"))) 
  
  #addCircles(clusterOptions = markerClusterOptions())

```


## Building Leaflet Map
```{r}

conn <- DBI::dbConnect(RPostgres::Postgres(), 
               dbname = 'diesel2021', 
               host='localhost',
               port='5432',
               user="cordero",
               password='Mojo')

metro <- st_read("./data/location_shapefiles/metro.shp") %>% st_transform(32610)

tier4_schools_query <- paste('
  SELECT p.*, st_value(r.rast, p.geometry) as "tier4_dpm", p.geometry
  FROM tier4_scenario AS r, schools AS p
  WHERE st_intersects(r.rast, p.geometry)
  ')

tier4_schools <- st_read(conn, query = tier4_schools_query)


## query for current scenario values
current_schools_query<- paste('
  SELECT p.*, st_value(r.rast, p.geometry) as "current_dpm", p.geometry
  FROM current_scenario AS r, schools AS p
  WHERE st_intersects(r.rast, p.geometry)
  ')

current_schools <- st_read(conn, query = current_schools_query)


## Filter NA values from current schools and tier 4 schools

current_schools <- current_schools %>% 
  filter(!is.na(current_dpm)) %>% 
  mutate(
    size = ifelse(current_dpm < 2, 2, current_dpm)
  )

tier4_schools <- tier4_schools %>% 
  filter(!is.na(tier4_dpm)) %>% 
  mutate(
    size = ifelse(tier4_dpm < 0.5, 0.5, tier4_dpm)
  )



tier4_scenario <- rpostgis::pgGetRast(conn, 'tier4_scenario')
current_scenario <- rpostgis::pgGetRast(conn, 'current_scenario')


current_schools_pal <- colorNumeric(
  palette = "YlOrRd",
  domain = current_schools$current_dpm
)

tier4_schools_pal <- colorNumeric(
  palette = "YlOrRd",
  domain = tier4_schools$current_dpm
)


tier4_scenario_pal <- colorNumeric("YlOrRd", values(tier4_scenario),
                    na.color = "transparent")
 
current_scenario_pal <- colorNumeric("YlOrRd", values(current_scenario),
                    na.color = "transparent")



leaflet() %>% 
  addProviderTiles(provider = "CartoDB.DarkMatter") %>% 
  ## add metro polygon with group id 'Portland Metro'
  addPolygons(data = st_transform(metro, 4326), color = "#b3b3b3", 
              fillOpacity = 0.1, weight = 2, 
              popup = "Portland Metro Area", group = 'Portland Metro', layerId = 'pdxmetro') %>% 
  # add Current Schools Exposure
  addCircleMarkers(data = st_transform(current_schools, 4326), radius = ~size, fillColor = ~current_schools_pal(current_dpm), group = 'Current School Exposure',
                   fillOpacity = 0.7, stroke = FALSE, 
                   clusterOptions = markerClusterOptions(), 
                   clusterId = "currentCluster", 
                   popup = ~paste("<b>School Name:</b>", name, 
                                  "<br><b>Current DPM Exposure:</b>", round(current_dpm, 2), "ppb",
                                  "<br><b>School Level:</b>", level_name,
                                  "<br><b>School Type:</b>", type,
                                  "<br><b>School Address:</b>", address, ",", state, zipcode)) %>% 
  # add Tier4 Schools Exposure
  addCircleMarkers(data = st_transform(tier4_schools, 4326), radius = ~5*size, fillColor = ~tier4_schools_pal(tier4_dpm), group = 'Tier 4 School Exposure',
                   fillOpacity = 0.7, stroke = FALSE, 
                   clusterOptions = markerClusterOptions(), 
                   clusterId = "tier4Cluster", 
                   popup = ~paste("<b>School Name:</b>", name, 
                                  "<br><b>Tier 4 DPM Exposure:</b>", round(tier4_dpm, 2), "ppb",
                                  "<br><b>School Level:</b>", level_name,
                                  "<br><b>School Type:</b>", type,
                                  "<br><b>School Address:</b>", address, ",", state, zipcode)) %>% 
  # add tier 4 scenario raster
  addRasterImage(tier4_scenario, colors = tier4_scenario_pal, opacity = 0.5, group = 'Tier 4 Scenario') %>%
  # add current scenario raster
  addRasterImage(current_scenario, colors = current_scenario_pal, opacity = 0.5, group = 'Current Scenario') %>%
  # Add Layers Control
  addLayersControl(overlayGroups = c("Current School Exposure", "Tier 4 School Exposure", "Current Scenario", "Tier 4 Scenario")) %>% 
  # Add Legends
  ## Tier 4 Scenario
  addLegend(pal = tier4_scenario_pal, values = values(tier4_scenario), title = "Tier 4 Scenario", position = "bottomright", group = "Tier 4 Scenario") %>% 
  ## Current Scenario
  addLegend(pal = current_scenario_pal, values = values(current_scenario), title = "Current Scenario", position = "bottomright", group = "Current Scenario") %>% 
  ## Current Schools Exposure
  addLegend(data = current_schools, pal = current_schools_pal, values = ~current_dpm, position = "bottomright", title = "Current DPM <br>Exposure (ppb)", group = "Current School Exposure") %>% 
  ## Tier 4 Schools Exposure
  addLegend(data = tier4_schools, pal = tier4_schools_pal, values = ~tier4_dpm, position = "bottomright", title = "Tier 4 DPM <br>Exposure (ppb)", group = "Tier 4 School Exposure") %>% 
   # Hide Layers/Legends at Start
  hideGroup(c("Tier 4 School Exposure", "Tier 4 Scenario", "Current Scenario")) %>% 
  addEasyButton(easyButton(
    states = list(
      easyButtonState(
        stateName="clustered-markers",
        icon="ion-toggle-filled",
        title="Remove Current Clusters",
        onClick = JS("
                     function(btn, map) {
                     var clusterManager =
                     map.layerManager.getLayer('cluster', 'currentCluster');
                     clusterManager.disableClustering();
                     btn.state('markers');
                     }")
        ),
  
      easyButtonState(
        stateName="markers",
        icon="ion-toggle",
        title="Enable Current Clusters",
        onClick = JS("
                     function(btn, map) {
                     var clusterManager =
                     map.layerManager.getLayer('cluster', 'currentCluster');
                     clusterManager.enableClustering();
                     btn.state('clustered-markers');
                     }")
         )
  )
  )) %>% 
  addEasyButton(easyButton(
    states = list(
      easyButtonState(
        stateName="clustered-markers",
        icon="ion-toggle-filled",
        title="Remove Tier 4 Clusters",
        onClick = JS("
                     function(btn, map) {
                     var clusterManager =
                     map.layerManager.getLayer('cluster', 'tier4Cluster');
                     clusterManager.disableClustering();
                     btn.state('markers');
                     }")
        ),
  
      easyButtonState(
        stateName="markers",
        icon="ion-toggle",
        title="Enable Tier 4 Clusters",
        onClick = JS("
                     function(btn, map) {
                     var clusterManager =
                     map.layerManager.getLayer('cluster', 'tier4Cluster');
                     clusterManager.enableClustering();
                     btn.state('clustered-markers');
                     }")
         )
  )
  )) %>% 
  addEasyButton(easyButton(
    icon="fa-crosshairs", title="Zoom To My Location",
    onClick=JS("function(btn, map){ map.locate({setView: true, maxZoom: 15}); }"))) %>% 
  addControl("PDX Schools Diesel PM Scenarios", position = "bottomleft")






```


